Discrete Quantum Theories and Computing

### 2018-02-27 dissertation.tex
Updates from ShareLaTeX
dissertation.lyx "Chapter 2 Conventional Quantum Theory" Updating...

\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\mathbb{F}_{{#1}}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\bf #1}}
\newcommand{\mathComplex}{\mathbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\mathbf{S}^{#1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\rmi}{i}
\newcommand{\rme}{e}

@@

Conventional quantum theory (CQT) is described by the following:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:-orthonormal-basis}$D$ orthonormal basis vectors for
a Hilbert space of dimension~$D$, 
\item \label{enu:probability-amplitude}the normalized $D$ complex probability
amplitude coefficients describing the contribution of each basis vector, 
\item \label{enu:unitary-matrix}a set of probability-conserving unitary
matrix operators that suffice to describe all required state transformations, 
\item \label{enu:measurement-framework}and a measurement framework. 
\end{enumerate}
We remark that there are many things that are assumed in CQT, such
as the absence of zero norm states for non-zero vectors, and the decomposition
of complex amplitudes into a pair of ordinary real numbers. One also
typically assumes the existence of a $D$-dimensional Hilbert space
with an orthonormal basis, allowing us to write \textit{pure} states
in general as Hilbert space vectors with an Hermitian inner product:
\begin{eqnarray}
\ket{\Psi} & = & \sum_{i=0}^{D-1}\alpha_{i}\ket{i}\,.
\end{eqnarray}
Here $\alpha_{i}\in\mathComplex$ are complex probability amplitudes,
$\VVec{\alpha}\in\mathComplex^{D}$, and the $\{\ket{i}\}$ is an
orthonormal basis of states obeying 
\begin{equation}
\braket{i}{k}=\delta_{ik}\,.
\end{equation}

The meaning of this is that any state $\ket{\Phi}=\sum_{i=0}^{D-1}\beta_{i}\ket{i}$
can be projected onto another state $\ket{\Psi}$ by writing 
\begin{eqnarray}
\braket{\Phi}{\Psi} & = & \sum_{i=0}^{D-1}\beta_{i}^{*}\alpha_{i}\,,
\end{eqnarray}
thus quantifying the proximity of the two states. (Here $^{*}$ denotes
complex conjugation.) This is one of many properties we take for granted
in continuum quantum mechanics that challenge us in defining a discrete
quantum geometry.

In Sec.~\ref{CQC1qubitBloch.sec}, we focus on the discrete geometric
issues raised by the properties \ref{enu:-orthonormal-basis} and
\ref{enu:probability-amplitude} given above for CQT. In Sec.~\ref{sec:Quantum-Circuit-Model},
we describe product and entangled $n$-qubit states used in circuit
model, and the unitary matrices used to transform states as addressed
by by the property \ref{enu:unitary-matrix}. In Sec.~\ref{sec:fuzzy},
we introduces the important issues of \ref{enu:measurement-framework}
and the foundations of quantum probability space. To facilitate the
transition to DQT carried out in later sections, we concern ourselves
first with the properties of the simplest possible abstract state
object in CQT, the single qubit state.

@@

@@

A state in a two-dimensional Hilbert space, also called a qubit, already
provides access to a wealth of geometric information and context.
We write the single qubit state as 
\begin{align}
\ket{\psi_{1}} & =\alpha_{0}\ket{0}+\alpha_{1}\ket{1} & \alpha_{0},\alpha_{1}\in\mathbb{C}\,.\label{A1.eq}
\end{align}

A convenience for probability calculations and a necessity for computing
relative state properties is the normalization condition 
\begin{equation}
\|\psi_{1}\|^{2}=|\alpha_{0}|^{2}+|\alpha_{1}|^{2}=\alpha_{0}^{*}\alpha_{0}+\alpha_{1}^{*}\alpha_{1}=1\ ,\label{A2.eq}
\end{equation}
which identifies $\alpha_{0}$ and $\alpha_{1}$ as (complex) probability
amplitudes and implies the conservation of probability in the closed
world spanned by $\{\ket{0},\ket{1}\}$. Note that we distinguish
for future use the \textit{norm\/} $\|\cdot\|$ of a vector from
the \textit{modulus\/} $|\cdot|$ of a complex number. Continuing,
we see that if we want only the irreducible state descriptions, we
must supplement the process of computing \Eq{A2.eq} by finding a
way to remove the distinction between states that differ only by an
overall phase transformation $\rme^{\rmi\phi}$, that is, 
\begin{equation}
\left(\alpha_{0},\alpha_{1}\right)\sim\left(\rme^{\rmi\phi}\alpha_{0},\rme^{\rmi\phi}\alpha_{1}\right)\,.\label{A3.eq}
\end{equation}
This can be accomplished by the Hopf fibration, which we can write
down as follows: let 
\begin{align}
\alpha_{0} & =x_{0}+\rmi y_{0}\,, & \alpha_{1} & =x_{1}+\rmi y_{1}\,.\label{A4.eq}
\end{align}
Then \Eq{A2.eq} becomes the condition that the four real variables
describing a qubit denote a point on the three-sphere $\Sphere{3}$
(a 3-manifold) embedded in $\R^{4}$: 
\begin{equation}
x_{0}^{\ 2}+y_{0}^{\ 2}+x_{1}^{\ 2}+y_{1}^{\ 2}=1\ .\label{A5.eq}
\end{equation}

>>>>

\newcommand{\muB}{\ensuremath{\mu^{B}}}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\gmult}{*}
\newcommand{\Fpx}[1]{\mathbb{F}_{{#1}}}
\newcommand{\Fq}{\Fpx{q}}
\newcommand{\ff}[1]{\Fpx{#1}}
\newcommand{\ffzx}[2]{\Fpx{#1}^{{#2}\;*}}
\newcommand{\ffx}[2]{\Fpx{#1}^{{#2}}}
\newcommand{\ffzd}[1]{\ffzx{#1}{d}}
\newcommand{\ffd}[1]{\ffx{#1}{d}}
\newcommand{\VVec}[1]{\vec{\bf #1}}
\newcommand{\mathComplex}{\mathbb{C}}
\newcommand{\braket}[2]{\ip{#1}{#2}}
\newcommand{\todo}[1]{\textbf{TODO.~#1}}
\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Sphere}[1]{\mathbf{S}^{#1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\rme}{\mathrm{e}}
\newcommand{\rmi}{\mathrm{i}}

@@

The part of conventional quantum theory (CQT) used by quantum circuit
model is described by the following:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:-orthonormal-basis}$D$ orthonormal basis vectors for
a Hilbert space of dimension~$D$, 
\item \label{enu:probability-amplitude}the normalized $D$ complex probability
amplitude coefficients describing the contribution of each basis vector, 
\item \label{enu:unitary-matrix}a set of probability-conserving unitary
matrix operators that suffice to describe all required state transformations, 
\item \label{enu:measurement-framework}and a measurement framework. 
\end{enumerate}
In Sec.~\ref{CQC1qubitBloch.sec}, we focus on the discrete geometric
issues raised by the properties \ref{enu:-orthonormal-basis} and
\ref{enu:probability-amplitude} given above for CQT. In Sec.~\ref{sec:Quantum-Circuit-Model},
we describe product and entangled $n$-qubit states used in quantum
circuit model, and the unitary matrices used to build circuit as addressed
by by the property \ref{enu:unitary-matrix}. In Sec.~\ref{sec:fuzzy},
we introduce the important issues of \ref{enu:measurement-framework}
and the foundations of quantum probability space.

@@

There are many things that are assumed in CQT, such as the absence
of zero norm states for non-zero vectors, and the decomposition of
complex amplitudes into a pair of ordinary real numbers. One also
typically assumes the existence of a $D$-dimensional Hilbert space
with an orthonormal basis, allowing us to write \textit{pure} states
in general as Hilbert space vectors with an Hermitian inner product:
\begin{eqnarray}
\ket{\Psi} & = & \sum_{i=0}^{D-1}\alpha_{i}\ket{i}\,.
\end{eqnarray}
Here $\alpha_{i}\in\mathComplex$ are complex probability amplitudes,
$\VVec{\alpha}\in\mathComplex^{D}$, and the $\{\ket{i}\}$ is an
orthonormal basis of states obeying 
\begin{equation}
\braket{i}{k}=\delta_{ik}\,.
\end{equation}

The meaning of this is that any state $\ket{\Phi}=\sum_{i=0}^{D-1}\beta_{i}\ket{i}$
can be projected onto another state $\ket{\Psi}$ by writing 
\begin{eqnarray}
\braket{\Phi}{\Psi} & = & \sum_{i=0}^{D-1}\beta_{i}^{*}\alpha_{i}\,,
\end{eqnarray}
thus quantifying the proximity of the two states. (Here $^{*}$ denotes
complex conjugation.) This is one of many properties we take for granted
in continuum quantum mechanics that challenge us in defining a discrete
quantum geometry. To facilitate the transition to DQT carried out
in later sections, we concern ourselves first with the properties
of the simplest possible abstract state object in CQT, the single
qubit state.

@@

A state in a two-dimensional Hilbert space, also called a qubit, already
provides access to a wealth of geometric information and context.
We write the single qubit state as 
\begin{align}
\ket{\psi_{1}} & =\alpha_{0}\ket{0}+\alpha_{1}\ket{1} & \alpha_{0},\alpha_{1} & \in\mathbb{C}\,.\label{A1.eq}
\end{align}

A convenience for probability calculations and a necessity for computing
relative state properties is the normalization condition 
\begin{equation}
\|\psi_{1}\|^{2}=|\alpha_{0}|^{2}+|\alpha_{1}|^{2}=\alpha_{0}^{*}\alpha_{0}+\alpha_{1}^{*}\alpha_{1}=1\ ,\label{A2.eq}
\end{equation}
which identifies $\alpha_{0}$ and $\alpha_{1}$ as (complex) probability
amplitudes and implies the conservation of probability in the closed
world spanned by $\{\ket{0},\ket{1}\}$. Note that we distinguish
for future use the \textit{norm\/} $\|\cdot\|$ of a vector from
the \textit{modulus\/} $|\cdot|$ of a complex number. Continuing,
we see that if we want only the irreducible state descriptions, we
must supplement the process of computing \Eq{A2.eq} by finding a
way to remove the distinction between states that differ only by an
overall phase transformation $\rme^{\rmi\phi}$, that is, 
\begin{equation}
\left(\alpha_{0},\alpha_{1}\right)\sim\left(\rme^{\rmi\phi}\alpha_{0},\rme^{\rmi\phi}\alpha_{1}\right)\,.\label{A3.eq}
\end{equation}
This can be accomplished by the Hopf fibration \cite{Hatcher2001,MosseriDandoloff2001,Bengtsson2007},
which we can write down as follows: let 
\begin{align}
\alpha_{0} & =x_{0}+\rmi y_{0}\,, & \alpha_{1} & =x_{1}+\rmi y_{1}\,.\label{A4.eq}
\end{align}
Then \Eq{A2.eq} becomes the condition that the four real variables
describing a qubit denote a point on the three-sphere $\Sphere{3}$
(a 3-manifold) embedded in $\R^{4}$: 
\begin{equation}
x_{0}^{\ 2}+y_{0}^{\ 2}+x_{1}^{\ 2}+y_{1}^{\ 2}=1\ .\label{A5.eq}
\end{equation}
